---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Gaynor et al paper 
    #https://faculty.wharton.upenn.edu/wp-content/uploads/2014/09/46_Competition_in_Health_Care_Markets.pdf
    
source("~/auth-aws.r")
Sys.setenv("AWS_DEFAULT_REGION" = "us-east-2")
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = FALSE)
source(here::here("R/manifest.r"))

# library(tidyverse)
# library(tidygraph)
# library(igraph)
# library(ggraph)
# library(ggdendro)
# library(hrbrthemes)
# library(patchwork)
# library(ggthemes)
# library(sf)
# library(glue)
# library(here)
# library(knitr)
set.seed(12345)
source(here("R/shared-functions.r"))
states <- c(
       "AL",    "AR",    "AZ",    "CA",    "CO",    "CT",    "DC",    "DE",    
    "FL",    "GA",      "IA",    "ID",    "IL",    "IN",    "KS",    "KY",    
    "LA",    "MA",    "MD",    "ME",    "MI",    "MN",    "MO",    "MS",    "MT",    
    "NC",    "ND",    "NE",    "NH",    "NJ",    "NM",    "NV",    "NY",    "OH",    
    "OK",    "OR",    "PA",    "RI",    "SC",    "SD",    "TN",    "TX",    "UT",    
    "VA",    "VT",    "WA",    "WI",    "WV",    "WY"
)
```

# A Basic Example

We'll start with a simple example. Here is the adjacency matrix representing a patient demand system. 

```{r}
B <-
    data.frame(A = c(rpois(2,lambda=10),rep(0,10)),
               B = c(rpois(3,lambda=15),rep(0,9)),
               C = c(rep(0,1),rpois(5,lambda=100),rep(0,6)),
               D = c(rep(0,2),rpois(4,lambda=50),rep(0,6)),
               E = c(rep(0,2),rpois(4,lambda=400),rep(0,6)),
               F = c(rep(0,2),rpois(4,lambda=100),rep(0,6)),
               G = c(rep(0,5),rpois(1,lambda=50), rpois(3,lambda=100),rep(0,1),rpois(2,lambda=500)),
               H = c(rep(0,6),rpois(5,lambda=100),0),
               I = c(rep(0,6),rpois(5,lambda=50),0),
               J = c(rep(0,7),rpois(4,lambda=200),0)
    ) %>%
    as.matrix()
colnames(B) <- paste0("HOSP_",colnames(B))
# force geo_8 to be a small monopoly from the perspective of the ZIP
B[8,] = c(rep(0,9),25)
rownames(B) <- paste0("GEO_",1:12)
```

```{r}
G <- get_bipartite_graph(B)
B_z <- B %>% create_unipartite_adjacency(type = "geo")
B_j <- B %>% create_unipartite_adjacency(type = "firm")
res <- B %>% analyze_market() 

# Analyze all potential pairwise mergers
mergers <- 
    expand.grid(colnames(B), colnames(B)) %>% 
    filter(Var1 != Var2) %>% 
    rowwise() %>% 
    mutate(pair = paste0(sort(c(Var1,Var2)),collapse="")) %>% 
    group_by(pair) %>% 
    filter(row_number()==1) %>% 
    ungroup() %>% 
    mutate(merge_id = row_number()) %>% 
    group_by(merge_id,Var1,Var2,pair) %>% 
    nest() %>% 
    mutate(res = map2(Var1,Var2,~({
        merged <- c(.x,.y)
        B_same <- B[,-which(colnames(B) %in% merged)] 
        B_new <- apply(B[,which(colnames(B) %in% merged)],1,sum)
        B_merged <- cbind(B_same,B_new)
        colnames(B_merged) <- c(colnames(B_same),paste0(merged,collapse ="_m_"))
        tmp <- B_merged %>% analyze_market()
        tmp$Y_j %>% bind_rows(tmp$Y_z)
    }))) %>% 
    unnest(cols=c(res))
```

```{r, echo = FALSE}
# Figure 1
fig_demand_toy <- B %>% ggmatrix()
ggsave(plot = fig_demand_toy, here("manuscript/tables-and-figures/figure-toy-demand-system-adjacency-bipartite.png"),
       height = 6, width = 6)
ggsave(plot = fig_demand_toy, here("manuscript/tables-and-figures/figure-toy-demand-system-adjacency-bipartite.pdf"),
       height = 6, width = 6)


```

# Market Detection

## Geographic Markets

```{r}

unipartite_z <- B_z %>% ggmatrix()
ggsave(plot =unipartite_z, here("manuscript/tables-and-figures/figure-toy-demand-system-adjacency-unipartite-geo.png"),
       height = 6, width = 6)
ggsave(plot = unipartite_z, here("manuscript/tables-and-figures/figure-toy-demand-system-adjacency-unipartite-geo.pdf"),
       height = 6, width = 6)
```

## Firm-Centered Markets

We begin by constructing the unipartite (firm-firm) matrix. 

```{r}
unipartite_j<- B_j %>% ggmatrix()

ggsave(plot =unipartite_j, here("manuscript/tables-and-figures/figure-toy-demand-system-adjacency-unipartite-firm.png"),
       height = 6, width = 6)
ggsave(plot = unipartite_j, here("manuscript/tables-and-figures/figure-toy-demand-system-adjacency-unipartite-firm.pdf"),
       height = 6, width = 6)
unipartite_j
```

The diagonal of this matrix is simply the total volume. The off-diagonal elements tell us, for a given hospital (row), the total number of its patients who are drawn from similar geographies as the paired hospital (column). 

Take the first row, for example. Cell [1,1] tells us Hospital A has a total volume of 23 patients. Cell [1,2] tells us that all 23 of Hospital A's patients are drawn from geographies from which Hospital B also draws its patients. By comparison, only 12 of Hospital A's patients are drawn from common geographies as Hospital C. 

We can verify all this by looking at the bipartite adjacency matrix. Hospital A draws 11 patients from GEO_1 and 12 from GEO_2. Hospital C's catchment area includes GEO_2, GEO_3, GEO_4, GEO_5, and GEO_6. There is one area of overlap (GEO_2) and from Hospital A's perspective 12 patients are drawn from this area. Alternatively, when pairing Hospital C and Hospital A (cell [3,1]) we see that 81 of Hospital C's 481 total patients are drawn from common geographies with Hospital A. 


```{r}
set.seed(123)
tmp <- res$market_j$G %>%  igraph::simplify(., remove.loops = TRUE)
E(tmp)$weight_rs <- scales::rescale(E(tmp)$weight,to=c(1,20))
E(tmp)$arrow.mode="-"
tmp %>% plot(layout = layout_with_fr,edge.width=E(tmp)$weight_rs)
# save as graph-network-example.pdf with dimensions 10x8in

```

```{r}
p_dendro <- 
    res$market_j$dendro  %>% ggdendrogram(rotate=TRUE) +
    scale_y_continuous(limits = c(0,res$market_j$max_height), breaks = seq(0,res$market_j$max_height,1)) +
    labs(x = "Market clustering level", 
         y = "Firm ID") 
p_dendro

ggsave(plot =p_dendro, here("manuscript/tables-and-figures/figure-toy-demand-system-dendro.png"),
       height = 6, width = 9)
ggsave(plot = p_dendro, here("manuscript/tables-and-figures/figure-toy-demand-system-dendro.pdf"),
       height = 6, width = 9)

``` 

```{r}

### JG TAKEUP FROM HERE : THESE RESULTS ARE NOW IN A STACKED DATA FRAME
p_hhi <- 
    res$market_j$hhi %>% 
    ggplot(aes(x = level, y = jitter(hhi)), guide = FALSE) + 
    #geom_point(aes(size = N, group = market, colour =factor(market)), guide =FALSE) +
    theme_ipsum() +
    ggsci::scale_color_aaas(name="") +
    theme(legend.position = 'none') +
    scale_y_continuous(limits = c(0,10000), breaks = seq(0,10000,2000)) +
    geom_line(data = res$hhi, aes(x = level, y = mean_hhi),lwd=1) +
    scale_x_continuous(limits = c(0,res$market_j$max_height), breaks = seq(0,res$market_j$max_height,1)) +
    labs(x = "Market clustering level", 
         y = "HHI") 
p_hhi
```

# Application

```{r}
application_year = "2017"
application_area = "TN"
area_denom = "state"  # The broadest area considered
#target_area = "county" # 
collapse_to_system_level = TRUE

create_application = TRUE
if (!file.exists(here(glue("output/adjacency-matrices/B-{application_area}")))) create_application =TRUE 
if (create_application) {
    source(here("R/01_construct-analytic-file.r"))
}


tmp <- read_rds(here::here(glue::glue("output/adjacency-matrices/B-county-{application_area}")))

systems <- tmp$hospitals %>% pull(system_id) %>% unique()
B <- tmp$G %>% get_collapsed_incidence_matrix(x = system_id)
B <- B[,intersect(colnames(B),systems)]
labs <- tmp$hospitals %>% select(sysname,system_id) %>% unique() %>% pull(sysname)
names(labs) <- tmp$hospitals %>% select(sysname,system_id) %>% unique() %>% pull(system_id)
colnames(B) <- labs[colnames(B)]

res <- B %>% analyze_market()

# Analyze all potential pairwise mergers

if (!file.exists(here(glue("output/mergers/G-merged-{application_area}-{application_year}.rds")))) {

    mergers <- 
        expand.grid(colnames(B), colnames(B)) %>% 
        filter(Var1 != Var2) %>% 
        rowwise() %>% 
        mutate(pair = paste0(sort(c(Var1,Var2)),collapse="")) %>% 
        group_by(pair) %>% 
        filter(row_number()==1) %>% 
        ungroup() %>% 
        mutate(id = row_number()) %>% 
        group_by(id,Var1,Var2,pair) %>% 
        nest() %>% 
        mutate(hhi = map2(Var1,Var2,~({
            merged <- c(.x,.y)
            B_same <- B[,-which(colnames(B) %in% merged)]
            B_new <- apply(B[,which(colnames(B) %in% merged)],1,sum)
            B_merged <- cbind(B_same,B_new)
            colnames(B_merged) <- c(colnames(B_same),paste0(merged,collapse ="_"))
            tmp <- B_merged %>% analyze_market()
            tmp$Y_j %>% bind_rows(tmp$Y_z)
        }))) %>% 
        unnest(cols=c(hhi))

    mergers %>% write_rds(here(glue("output/mergers/G-merged-{application_area}-{application_year}.rds")))
} else {
  mergers  <- read_rds(here(glue("output/mergers/G-merged-{application_area}-{application_year}.rds")))  
}


p_dendro <- 
res$market_j$dendro  %>% ggdendrogram(rotate=TRUE) +
scale_y_continuous(limits = c(0,res$market_j$max_height), breaks = seq(0,res$market_j$max_height,1)) +
labs(x = "Market clustering level", 
     y = "Firm ID") +
scale_y_continuous(limits = c(0,res$market_j$max_height*1.2), breaks = scales::breaks_pretty())  +
labs(y = "Clustering level")


# res$Y_j %>% ungroup() %>%  select(level,modularity) %>% unique() %>% 
#     ggplot(aes(x = level, y= modularity))+ geom_point() + 
#     theme_ipsum()

# p_dendro + geom_hline(aes(yintercept = res$Y_j %>% filter(max_modularity==1) %>% pull(level) %>% unique() ), 
#                       colour = "red")

p_hhi <- 
res$market_j$hhi %>% 
ggplot(aes(x = level, y = jitter(hhi)), guide = FALSE) + 
#geom_point(aes(size = N, group = market, colour =factor(market)), guide =FALSE) +
theme_ipsum() +
ggsci::scale_color_aaas(name="") +
theme(legend.position = 'none') +
scale_y_continuous(limits = c(0,10000), breaks = seq(0,10000,2000)) +
geom_line(data = res$hhi, aes(x = level, y = mean_hhi),lwd=1) +
scale_x_continuous(limits = c(0,res$market_j$max_height), breaks = seq(0,res$market_j$max_height,1)) +
labs(x = "Market clustering level", 
     y = "HHI") 
```

```{r, fig.height=8,fig.width=8}
set.seed(123)

tmp <- res$market_j$G %>%  igraph::simplify(., remove.loops = TRUE) %>% 
    as_tbl_graph() %>% 
    activate(edges) %>%
    filter(weight>5000) %>% 
    as.igraph()
E(tmp)$weight_rs <- scales::rescale(E(tmp)$weight,to=c(0,2))
E(tmp)$arrow.mode="-"
V(tmp)$size = 1
V(tmp)$label.color <- "black"
#V(tmp)$label <- NA
#tmp %>% plot(layout =layout_in_circle,edge.width=E(tmp)$weight_rs, vertex.label.cex=.5)
tmp %>% plot(layout =layout_with_lgl,edge.width=E(tmp)$weight_rs, vertex.label.cex=.5)
```

```{r, fig.height = 10, fig.width =10}
p_frontier <- 
    p_hhi +
    geom_line(data = mergers, aes(x = level, y = mean_hhi, group = id),alpha=0) 

df_frontier_hi <- 
    ggplot_build(p_frontier)$data[[2]] %>% 
    as_tibble() %>% 
    group_by(x) %>% 
    filter(y==max(y)) %>% 
    #filter(row_number()==1) %>% 
    select(level = x, mean_hhi = y ,id=group) %>% 
    inner_join(mergers,c("id","mean_hhi","level")) %>% 
    arrange(level,id) %>% 
    select(level,pair, mean_hhi)

df_frontier_lo <- 
    ggplot_build(p_frontier)$data[[2]] %>% 
    as_tibble() %>% 
    group_by(x) %>% 
    filter(y==min(y)) %>% 
    #filter(row_number()==1) %>% 
    select(level = x, mean_hhi = y ,id=group) %>% 
    inner_join(mergers,c("id","mean_hhi","level")) %>% 
    arrange(level,id) %>% 
    select(level,pair, mean_hhi)
hhi_all <- sum((100*apply(B,2,sum)/sum(apply(B,2,sum)))^2)

df_annotate <- 
    data.frame(x = c(0,res$market_j$max_height), y = c(10000,res$market_j$hhi    %>% filter(level==max(level)) %>% pull(hhi))) %>% 
    mutate(label = c("Each hospital\nis its own market (HHI=10,000)",glue("All hospitals in\none market (HHI={round(hhi_all,0)})")))

#km_hhi <- km_system %>% summarise(mean = weighted.mean(hhi_j,w=volume)) %>% pull(mean) %>% round(.,0)

p_frontier_final <- 
    p_frontier +
    geom_line(data =df_frontier_hi, aes(x=level, y = mean_hhi), colour ="red", alpha = 0.5, lty=1,lwd=1) +
    geom_line(data =df_frontier_lo, aes(x=level, y = mean_hhi), colour ="blue", alpha = 0.5, lty=1,lwd=1) +
    theme(panel.grid.major=element_blank(),
          panel.grid.minor=element_blank()) +
    geom_text(data = df_annotate, aes(x=x,y=y,label=label),size=3,hjust=-.2) +
    scale_x_continuous(limits = c(0,res$market_j$max_height*1.3), breaks = scales::breaks_pretty())  +
    geom_point(data = df_annotate, aes(x=x,y=y),size=3) #+ 
    #geom_hline(aes(yintercept = km_hhi),lty=3) #+
# annotate("text",x=res$market_j$max_height,y=km_hhi,
#          label=glue("Kessler-McClellan HHI: {km_hhi}"),size=4,vjust=-1) #+
    #geom_hline(aes(yintercept = hhi_all),lty=3) 

```

```{r}
data.frame(system = colnames(B)) %>% 
    as_tibble() %>% 
    rowwise() %>% 
    mutate(frequency = length(grep(system,df_frontier_hi %>% 
                                       filter(level>3) %>% 
    pull(pair) ))) %>% 
    mutate(frequency = 100* frequency / res$market_j$max_height) %>% 
    filter(frequency>0) %>%     
    arrange(desc(frequency)) %>% 
    mutate(frequency = glue("{round(frequency,1)}%")) %>% 
    kable(col.names=c("Hospital System", "Frequency on Frontier"))
```

# Market Outcome Measures

## HHI

### Outflow HHI

### Kessler-McClellan

```{r}
#df_hhi <- read_rds(here(glue("output/hhi/hhi-{application_area}.rds")))

# # Analyze current market
# km_system <- 
#     read_rds(here("output/results/kessler-mcclellan-2017-edgelist/km-2017-edgelist-system.rds"))$hosp %>% 
#     filter(system_id %in% intersect(colnames(B),systems)) %>% 
#     mutate(volume = apply(B,2,sum)[sysname]) 


```

